Stage-Splitting a Modal Language

### 2014-08-27 paper/semantics.tex
Added section on value splitting and lots of exmaples.

At all of the normal features (\ref{fig:diaSemanticsSpec}), speculation does
nothing but recursively speculate into every subexpression.  Once speculation
finds a $\prev$ block, we resume stage \bbone\ evaluation of the contents, which
produces a future context and (by canonical forms) a $\next$-wrapped variable
($\next\{\hat y\}$).  The context is then reified (using the $\reifysym$
judgment) into a series of let bindings, with $\mathtt{\hat y}$---stripped of
its $\next$---at the bottom.

>>>>

At all of the normal features (\ref{fig:diaSemanticsSpec}), speculation does
nothing but recursively speculate into every subexpression.  Once speculation
finds a $\prev$ block, we resume stage \bbone\ evaluation of the contents, which
produces a future context and (by canonical forms) a $\next$-wrapped variable
($\next\{\hat y\}$)\footnote{
Importantly, speculation will search into both branches of a stage-\bbtwo\ {\tt if}
statement to evaluate the contained stage-\bbone\ code.  Thus the name.
}.  The context is then reified (using the $\reifysym$
judgment) into a series of let bindings, with $\mathtt{\hat y}$---stripped of
its $\next$---at the bottom.

### 2014-08-27 paper/splitting.tex
Added section on value splitting and lots of exmaples.

The basic idea is that we want to take a term which contains interleaved stage \bbone\ and stage \bbtwo\ code
and split it into two separate terms: one with all of the stage \bbone\ code and one with all the stage \bbtwo\ code.
To facilitate communication between the stages, there is a data structure passed between the stages, 
which is an output of the stage \bbone\ term and and input to the stage \bbtwo\ term.

As with the dynamics, the form of the splitting statement depends on the external stage of the term.
We consider splitting for stage \bbtwo\ terms first, as it has a simpler form.  
For example, take \verb|hold{1+2} < 5|.
We can split this into two separate terms, \verb|1+2| and \verb|fn l => l < 5|, 
which are called the {\em precomputation} and {\em residual}
\footnote{We also used the term {\em residual} for the result of partial evaluation.
Where necessary, we will distinguish between these concepts by using the terms {\em splitting-residual}
and {\em evaluation-residual}.}, respectively.

@@

@@

\subsection {Stage \bbone\ Divergence}
Consider the case of stage \bbone\ {\tt if} and {\tt case} expressions.
In both cases we have enough information at stage \bbone\ to know what branch to take, so there's no need to speculate.
Instead, we evaluate the stage \bbone\ portion of only the active branch, and then inject precomputation into a sum type.
Then in the residual, we case on that sum and resume the stage \bbtwo\ portion of the correct branch.

@@

Thus the boundary type appears as output of the first stage and input of the second stage.
This makes it hard to type the output, because the boundary types are not represented in the original function type.

\subsection {Boundary Type Worst Case}

\TODO add the example which is the worst case for figuring out boundary types

@@

\subsection{Value Splitting, Etc}

...

>>>>

The basic idea is that we want to take a \lang\ term which contains interleaved stage \bbone\ and stage \bbtwo\ code
and split it into two separate terms: one with all of the stage \bbone\ code and one with all the stage \bbtwo\ code.
To facilitate communication between the stages, there is a data structure passed between the stages, 
which is an output of the stage \bbone\ term and and input to the stage \bbtwo\ term.

As with the dynamics, the form of the splitting statement depends on the external stage of the term.
We consider splitting for stage \bbtwo\ terms first, as it has a simpler form.  
For example, take \verb|hold{1+2} < 5|.
We can split this into two separate terms, \verb|1+2| and \verb|fn l => l < 5|, 
which are called the {\em precomputation} and {\em residual}\footnote{We also used the term {\em residual} for the result of partial evaluation.
Where necessary, we will distinguish between these concepts by using the terms {\em splitting-residual}
and {\em evaluation-residual}.}, respectively.

@@

\subsection{Value Splitting}

Before getting into the formal presentation, 
we first consider the problem of splitting stage-\bbone\ partial values.
This serves as a didactic stepping-stone to full term-splitting, 
and it is a necessary component of the correctness theorem.

The rules for value splitting are given in \cref{fig:valSplit}.  It consists of a single judgment, $v \vsplito [v_1,v_2]$,
which sends a partial value $v$ to two single-stage values $v_1$ and $v_2$. 
Those respectively represent the stage-\bbone\ and stage-\bbtwo\ content of the original value.  
At base types, value splitting assigns all content to stage \bbone, and puts in a nullary tuple for stage two,
in accordance with the precept that base types have only trivial content at stage \bbtwo.
Splitting essentially distributes into values at product, sum, and recursive types.
Functions split into two other functions.  Since the body of a function is an open term, this rule must rely on 
general term splitting, which is covered momentarily.

Lastly, we consider value splitting at the $\fut$ type.  The only partial value with this type
is a $\next$ block containing a reference into the future environment.  This splits into a 
nullary tuple and the same variable, again aligning with the precept that $\fut$ types contain trivial stage-\bbone\
content.

@@

\subsection {Stage \bbone\ Divergence}
Consider the splitting rules for stage-\bbone\ {\tt if} and {\tt case} expressions.
In both cases we have enough information at stage \bbone\ to know what branch to take, so there's no need to speculate.
Instead, we evaluate the stage \bbone\ portion of only the active branch, and then inject precomputation into a sum type.
Then in the residual, we case on that sum and resume the stage \bbtwo\ portion of the correct branch.

@@

%\subsection {Boundary Type Worst Case}
%
%\TODO add the example which is the worst case for figuring out boundary types
%

@@
